### 2019-05-19 chapter_00_open_science.tex
Ch 0: Rewrite intro

% \subsection{Science, software, open-source and the computer revolution}

@@

This chapter presents FluidDyn, a project and an ecosystem of packages to
facilitate open-science and open-source development in the field of fluid
mechanics.
%
The chapter is organized into several parts to present the context and to
describe our motivations and the concept behind the project FluidDyn as
follows:

\begin{itemize}
\item A scientific project based on the collaboration through open-source, such
as FluidDyn, is only possible thanks to the recent developments in methods and
tools for open-source software engineering. We start by presenting these
developments and discuss the implications for scientists.

\item FluidDyn is based on the language Python. Incidentally, it is the defining
characteristics of Python which make the eventual success of such a
project realizable.
%
Python is one of the most important tool in recent open-source dynamics,
particularly in science.
%
Python in fluid mechanics is growing fast, but is yet to reach widespread
adoption as the language of choice. Therefore, it is useful to present its
characteristics for readers who are unacquainted with its nuances.

\item We will discuss about the possible contradiction between productivity of
individuals and productivity at the community level.

\item We will establish a picture of software engineering in fluid mechanics
and present some of the main tools commonly used.

\item The introduction finally ends with a summary of the motivations for the
project FluidDyn and an overview of the packages used\footnote{%
  We use FluidDyn (with capital letters) to name the project and \fluiddyn for
the base package.}.
\end{itemize}

\section{A modern approach to computational sciences}

>>>>

@@

In order to attain the research goals in this thesis, a path off the beaten
track was pursued -- i.e. to create a handful of easy to maintain, rigorously
tested, and most-importantly, open-source scientific packages which does not
compromise on performance.
%
This resulted in the creation of FluidDyn, a project to facilitate open-science
in the field of fluid mechanics.
%
The focus of the project was to develop a set of packages to implement a
framework for simulations \citep{fluidsim}, experiments
\citep{augier_fluidimage_2016,augier_fluidlab_nodate}, Fast Fourier Transforms
\citep{fluidfft} -- all depending on a base package to reuse certain utilities
and data structures \citep{fluiddyn}.
%
The project owes a lot to the recent developments in methods and tools for
open-source software engineering.
%
Several good and modern coding practices were adopted while developing these
packages.
%
An understanding of the design choices and the workflow involved would be of
general interest to the reader.

FluidDyn packages are primarily written in Python language. Incidentally, it is
the defining characteristics of Python which make the eventual success of such
a project realizable.
%
Python is one of the most important tool in recent open-source dynamics,
particularly in science.
%
The popularity of the language in fluid mechanics is growing, but is yet
to reach widespread adoption as the language of choice.

We shall now discuss these topics in greater detail. The chapter is organized
into two sections. The first section focusses on the technical aspects, with
particular focus on emerging developments in programming languages, tools and
services suited for open-source software development. We shall present Python's
strengths and weaknesses as a language for scientific computing.
%
The section ends with a note on the capabilities of FluidDyn
packages\footnote{We use FluidDyn (with capital letters) to name the project
and \fluiddyn for the base package.}.
%
In the second section a comparison of various software development
methodologies are made. Here, we will also discuss about the possible
contradiction between productivity of individuals and productivity at the
community level.
%
% The chapter finally ends with a summary of the motivations for the project
% FluidDyn and an overview of the packages used.

\section{Science, software, open-source and the computer revolution}
